### 2020-04-13 serial.tex
we
reorder

@@

\fref{rR} compares the estimates of the reproduction number $\mathcal R$ from the exponential growth rate $r$ using forward serial-interval distributions and naive serial-interval distributions that do not account for the differences between backward and forward incubation periods.
When we use the forward serial-interval distributions to estimate $\mathcal R$, we obtain identical values regardless of the correlation between incubation periods and generation intervals;
as expected, these estimates match $\mathcal R$ estimes using the intrinsic generation-interval distributions (\fref{rR}A).
On the other hand, when naive serial-interval distributions are used, we underestimate $\mathcal R$; $\mathcal R$ even decreases for high $r$ because negative serial intervals are given too much weight (\fref{rR}B).
While the forward serial intervals can be also negative, the proportion of negative intervals are appropriately balances because faster epidemic growth will lead to shorter backward incubation period (and therefore less negative serial intervals).

@@

Here, we provide a simple, heuristic way of assessing potential biases in the estimate of $\mathcal R$ retrospctively.
Once a set of serial intervals has been observed from the line list, we can group observed serial intervals by their primary cohort times.
Then, we can compare how estimates of $\mathcal R$ change as we include more cohorts into the analysis (see `cohort-averaged' in \fref{obsrR}).
During the exponential growth phase, the estimates of $\mathcal R$ remains consistent;
adding more data allows us to make more precise inference about $\mathcal R$ during this period.
However, the cohort-averaged estimates of $\mathcal R$ decreases rapidly soon after the exponential growth period.
Confidence intervals associated with the cohort-averaged estimates of $\mathcal R$ becomes narrower and do not overlap with earlier estimates.
This approach allows us to visualize how changes in the forward serial intervals translate to biases in estimates of $\mathcal R$.

>>>>

First, we compare the estimates of the reproduction number $\mathcal R$ based on the intrinsic generation-interval distribution and the forward serial-interval distribution during the exponential growth phase (\fref{rR}A).
Both distributions provide identical estimates of $\mathcal R$ regardless of the correlation $\rho$ between the incubation period distribution and the intrinsic generation-interval distribution.
On the other hand, using naive serial-interval distributions that do not account for disease dynamics (i.e., assuming that the backward and the forward incubation period distributions are identical) underestimates $\mathcal R$;
as $r$ increases, estimates of $\mathcal R$ eventually saturate and decrease because negative serial intervals are given too much weight (\fref{rR}B)
While the forward serial intervals during the exponential growth phase can be also negative, the proportion of negative intervals are appropriately balanced because faster epidemic growth will lead to shorter backward incubation period (and therefore less negative serial intervals).

@@

@@

We provide a simple, heuristic way of assessing potential biases in the estimate of $\mathcal R$ retrospctively.
Once a set of serial intervals has been observed from the line list, we can group observed serial intervals by their primary cohort times.
Then, we can compare how estimates of $\mathcal R$ change as we include more cohorts into the analysis (see `cohort-averaged' in \fref{obsrR}).
During the exponential growth phase, the estimates of $\mathcal R$ remains consistent;
adding more data allows us to make more precise inference about $\mathcal R$ during this period.
However, the cohort-averaged estimates of $\mathcal R$ decreases rapidly soon after the exponential growth period.
Confidence intervals associated with the cohort-averaged estimates of $\mathcal R$ becomes narrower and do not overlap with earlier estimates.
This approach allows us to visualize how changes in the forward serial intervals translate to biases in estimates of $\mathcal R$.

Finally, we apply these ideas and reanalyze serial intervals of COVID-19 collected from mainland China, outside Hubei province, between XXX and XXX (REF).
